---
title: "MSI Compliance Analysis"
author: "Analysis by Rachel Spicer, github:RASpicer"
date: "2017/08/05"
output: 
  html_document:
    code_folding: hide
    number_sections: yes
    theme: cerulean
    keep_md: true
---

```{r setup, include=FALSE}
list.of.packages <- c("ggplot2","knitr","xlsx","FSA")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
# install.packages(list.of.packages) # run this if some packages are too old and cause errors
# from http://stackoverflow.com/questions/4090169/elegant-way-to-check-for-missing-packages-and-install-them

library(ggplot2)
library(xlsx)
library(knitr)
library(FSA)

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, tidy=TRUE, message = FALSE,fig.path = "figs/",fig.keep=TRUE)

```

# Data Processing
This RMarkdown contains the code used for analysis for the 2017 paper <b>Compliance with Minimum Information Guidelines in Public Metabolomics Repositories</b>.

## Functions

This section includes code for extracting only the minimal or best practice/ optional reporting standards.

```{r functions}
# Calculate the percentage compliance with the minimal reporting standards for clinical data
MinPerClin <- function(x) {
  # Calculate totals
  Totals = as.data.frame(colSums(x, na.rm = TRUE))
  # Extract only minimal reporting standards
  MinRep = c(Totals[1:7,], Totals[14:22,], Totals[42:44,], Totals[46,], Totals[48,], Totals[55,])
  # Calculate percentage
  Per = MinRep/(nrow(x)) * 100
  Per = round(Per, digits = 2)
  # Get column names
  Names = c(colnames(x[,1:7]), colnames(x[,14:22]), colnames(x[,42:44]), colnames(x[46]), colnames(x[48]), colnames(x[55]))
  MinMeta = as.data.frame(cbind(Names, Per))
  colnames(MinMeta) = c("Metadata", "Percentage")
  MinMeta$Percentage = as.numeric(as.character(MinMeta$Percentage))
  # Return percentage
  return(MinMeta)
}

# Calculate the percentage compliance with the minimal reporting standards for preclinical data
MinPerPreClin <- function(x,y) {
  # Calculate totals
  Totals = as.data.frame(colSums(x, na.rm = TRUE))
  # Extract only minimal reporting standards
  MinRep = c(Totals[1,], Totals[3:7,], Totals[14:17,], Totals[22:23,], Totals[25:30,], Totals[32:34,], Totals[36:37,], Totals[39:45,])
  # Calculate percentage
  Per = MinRep/(nrow(x)) * 100
  Per = round(Per, digits = 2)
  # Get column names
  Names = c(colnames(x[1]), colnames(x[,3:7]), colnames(x[,14:17]), colnames(x[,22:23]), colnames(x[,25:30]), colnames(x[,32:34]), colnames(x[,36:37]), colnames(x[,39:45]))
  MinMeta = as.data.frame(cbind(Names, Per))
  colnames(MinMeta) = c("Metadata", "Percentage")
  MinMeta$Percentage = as.numeric(as.character(MinMeta$Percentage))
  # Return percentage
  return(MinMeta)
}

# Calculate the percentage compliance with the minimal reporting standards for microbial/cell line data
MinPerCellLine <- function(x) {
  # Calculate totals
  Totals = as.data.frame(colSums(x, na.rm = TRUE))
  # Extract only minimal reporting standards
  MinRepSt = Totals[1:15,]
  # Calculate percentage
  Per = MinRepSt/(nrow(x)) * 100
  Per = round(Per, digits = 2)
  # Get column names
  Names = c(colnames(x[,1:15]))
  MinMeta = as.data.frame(cbind(Names, Per))
  colnames(MinMeta) = c("Metadata", "Percentage")
  MinMeta$Percentage = as.numeric(as.character(MinMeta$Percentage))
  return(MinMeta)
}

# Calculate the percentage compliance with the minimal reporting standards for plant studies
MinPerPlant <- function(x) {
  # Calculate totals
  Totals = as.data.frame(colSums(x, na.rm = TRUE))
  # Calculate percentage
  Per = Totals/(nrow(x)) * 100
  Per = round(Per, digits = 2)
  # Get column names
  Names = colnames(x)
  MinMeta = as.data.frame(cbind(Names, Per))
  colnames(MinMeta) = c("Metadata", "Percentage")
  MinMeta$Percentage = as.numeric(as.character(MinMeta$Percentage))
  # Return percentage
  return(MinMeta)
}

# Calculate the percentage compliance with the optional reporting standards for microbial/cell line data
OpPerCellLine <- function(x) {
  # Calculate totals
  Totals = as.data.frame(colSums(x, na.rm = TRUE))
  # Extract only optional reporting standards
  OpRep = Totals[16:54,]
  # Calculate percentage
  Per = OpRep/(nrow(x)) * 100
  Per = round(Per, digits = 2)
  # Get column names
  Names = c(colnames(x[,16:54]))
  OpMeta = as.data.frame(cbind(Names, Per))
  colnames(OpMeta) = c("Metadata", "Percentage")
  OpMeta$Percentage = as.numeric(as.character(OpMeta$Percentage))
  # Return percentage
  return(OpMeta)
}

# Calculate the percentage compliance with the optional reporting standards for clinical data
OpPerClin <- function(x) {
  # Calculate totals
  Totals = as.data.frame(colSums(x, na.rm = TRUE))
  OpRep = c(Totals[8:13,], Totals[23:41,], Totals[45,], Totals[47,], Totals[49:54,])
  # Calculate percentage
  Per = OpRep/(nrow(x)) * 100
  Per = round(Per, digits = 2)
  # Get column names
  Names = c(colnames(x[,8:13]), colnames(x[,23:41]), colnames(x[45]), colnames(x[47]), colnames(x[,49:54]))
  OpMeta = as.data.frame(cbind(Names, Per))
  colnames(OpMeta) = c("Metadata", "Percentage")
  OpMeta$Percentage = as.numeric(as.character(OpMeta$Percentage))
  # Return percentage
  return(OpMeta)
}

# Calculate the percentage compliance with the optional reporting standards for preclinical data
OpPerPreClin <- function(x) {
  # Calculate totals
  Totals = as.data.frame(colSums(x, na.rm = TRUE))
  # Extract only minimal reporting standards
  OpRep = c(Totals[2,], Totals[8:13,], Totals[18:21,], Totals[24,], Totals[31,], Totals[35,], Totals[38,], Totals[46,])
  # Calculate percentage
  Per = OpRep/(nrow(x)) * 100
    Per = round(Per, digits = 2)
  # Get column names
  Names = c(colnames(x[2]), colnames(x[,8:13]), colnames(x[,18:21]),colnames(x[24]), colnames(x[31]), colnames(x[35]), colnames(x[38]), colnames(x[46]))
  OpMeta = as.data.frame(cbind(Names, Per))
  colnames(OpMeta) = c("Metadata", "Percentage")
  OpMeta$Percentage = as.numeric(as.character(OpMeta$Percentage))
  # Return percentage
  return(OpMeta)
}

# ggplot theme
theme_adjbw <- function(base_size = 12, base_family = ""){
  theme_bw() +
    theme(
      axis.text = element_text(colour = "black"),
      axis.line.x = element_line(color="black", size = 0.5),
      axis.line.y = element_line(color="black", size = 0.5),
      # Remove gridlines and borders
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.border = element_blank(),
      panel.background = element_blank()
    )
}

```

## Data Processing

Data for the compliance to one of the four sets of reporting standards is first split by repository. The percentage of studies that comply with each reporting standard is then calculated. Reporting standards are then divided into minimal and best practice/optional.

```{r data_processing}
# read compliance data for each study type
Mouse <- read.xlsx("../MousePreclinicalRawData4Analysis.xlsx", sheetIndex = 1, check.names = FALSE)
Plant <- read.xlsx("../ArabidopsisPlantRawData4Analysis.xlsx", sheetIndex = 1, check.names = FALSE)
HumanClin <- read.xlsx("../HumanClinicalRawData4Analysis.xlsx", sheetIndex = 1, check.names = FALSE)
HumanCell <- read.xlsx("../HumanCellRawData4Analysis.xlsx", sheetIndex = 1, check.names = FALSE)

# Extract only MetaboLights studies 
MLMouse <- Mouse[grep("MetaboLights", Mouse$`Study Repository `),]
MLPlant <- Plant[grep("MetaboLights", Plant$`Study Repository `),]
MLHumanClin <- HumanClin[grep("MetaboLights", HumanClin$`Study Repository`),]
MLHumanCell <- HumanCell[grep("MetaboLights", HumanCell$`Study Repository `),]

# Extract only Metabolomics Workbench studies
MWMouse <- Mouse[grep("Metabolomics Workbench", Mouse$`Study Repository `),]
MWPlant <- Plant[grep("Metabolomics Workbench", Plant$`Study Repository `),]
MWHumanClin <- HumanClin[grep("Metabolomics Workbench", HumanClin$`Study Repository`),]
MWHumanCell <- HumanCell[grep("Metabolomics Workbench", HumanCell$`Study Repository `),]

# Extract only MetaPhen studies
MPPlant <- Plant[grep("MetaPhen", Plant$`Study Repository `),]

# Extract only MeRy-B studies
MeRPlant <- Plant[grep("MeRy-B", Plant$`Study Repository `),]

# Minimal - MetaboLights
MLHumanCellP <- MinPerCellLine(MLHumanCell[,-(1:2)])
MLHumanCellP$Order <- rep(1)
MLHumanClinP <- MinPerClin(MLHumanClin[,-(1:2)])
MLHumanClinP$Order <- rep(2)
MLMouseP <- MinPerPreClin(MLMouse[,-(1:2)])
MLMouseP$Order <- rep(3)
MLPlantP <- MinPerPlant(MLPlant[,-(1:2)])
MLPlantP$Order <- rep(4)

# Minimal - Metabolomics Workbench
MWHumanCellP <- MinPerCellLine(MWHumanCell[,-(1:2)])
MWHumanCellP$Order <- rep(1)
MWHumanClinP <- MinPerClin(MWHumanClin[,-(1:2)])
MWHumanClinP$Order <- rep(2)
MWMouseP <- MinPerPreClin(MWMouse[,-(1:2)])
MWMouseP$Order <- rep(3)
MWPlantP <- MinPerPlant(MWPlant[,-(1:2)])
MWPlantP$Order <- rep(4)

# Minimal - MetaPhen
MPPlantP <- MinPerPlant(MPPlant[,-(1:2)])
MPPlantP$Order <- rep(1)

# Minimal - MeRy-B
MeRPlantP <- MinPerPlant(MeRPlant[,-(1:2)])
MeRPlantP$Order<- rep(4)

# Best Practice - MetaboLights
MLHumanCellOP <- OpPerCellLine(MLHumanCell[,-(1:2)])
MLHumanCellOP$Order <- rep(1)
MLHumanClinOP <- OpPerClin(MLHumanClin[,-(1:2)])
MLHumanClinOP$Order <- rep(2)
MLMouseOP <- OpPerPreClin(MLMouse[,-(1:2)])
MLMouseOP$Order <- rep(3)

# Best Practice - Metabolomics Workbench
MWHumanCellOP <- OpPerCellLine(MWHumanCell[,-(1:2)])
MWHumanCellOP$Order <- rep(1)
MWHumanClinOP <- OpPerClin(MWHumanClin[,-(1:2)])
MWHumanClinOP$Order <- rep(2)
MWMouseOP <- OpPerPreClin(MWMouse[,-(1:2)])
MWMouseOP$Order <- rep(3)
```

# Figures
Code that was used to generate the raw figures from the paper. The raw figures were then edited in Adobe Illustrator to add letters to indicate statistical significance and to change colours to patterns in figure 2.

## Figure 1a. Compliance to Minimal MSI Reporting Standards in MetaboLights
Combined box-and-whisker and dot plot showing the percentage compliance with the MSI minimum reporting standards within the MetaboLights repository. and b) Metabolomics Workbench repositories. Each “dot” represents compliance to a single reporting standard.

```{r MLMin, results = 'asis', fig.height = 4.5, fig.width = 4, fig.align = "center"}
ML <- rbind(MLHumanCellP, MLHumanClinP, MLMouseP, MLPlantP)
ML$Order <- as.factor(ML$Order)
ggplot(ML, aes(x=Order, y=Percentage))  + 
  geom_boxplot(fill = "#7fc97f", outlier.shape = 3, outlier.color = "red")  +
  geom_jitter() +
  scale_x_discrete(labels = c(expression(italic("In vitro")),
                              "Clinical",
                              "Pre-clinical",
                              "Plant")) +
  scale_y_continuous(expand = c(0, 0), limits = c(-1, 101)) +
  theme_adjbw() +
  xlab("MSI Guideline") +
  ylab("Compliance (%)")
```

## Figure 1b. Compliance to Minimal MSI Reporting Standards in Metabolomics Workbench
Combined box-and-whisker and dot plot showing the percentage compliance with the MSI minimum reporting standards within the Metabolomics Workbench repository. Red ‘+’ indicate outliers, each “dot” represents compliance to a single reporting standard.

```{r MWMin, results = 'asis', fig.height = 4.5, fig.width = 4, fig.align = "center"}
MW <- rbind(MWHumanCellP, MWHumanClinP, MWMouseP, MWPlantP)
MW$Order <- as.factor(MW$Order)
ggplot(MW, aes(x=Order, y=Percentage))  + 
  geom_boxplot(fill = "#beaed4", outlier.shape = 3, outlier.color = "red")  +
  geom_jitter() +
  scale_x_discrete(labels = c(expression(italic("In vitro")),
                              "Clinical",
                              "Pre-clinical",
                              "Plant")) +
  theme_adjbw() +
  scale_y_continuous(expand = c(0, 0), limits = c(-1, 101)) +
  xlab("MSI Guideline") +
  ylab("Compliance (%)")
```

## Figure 2a. Comparison of Compliance to Minimal and Best Practice Reporting Standards in MetaboLights
Box-and-whisker plot showing the percentage compliance to the minimal and optional/best practice reporting standards in MetaboLights. Minimal reporting standards are indicated in green, optional/best practice reporting standards are blue and red ‘+’ indicate outliers.

```{r MLCom, results = 'asis', fig.height = 4.5, fig.width = 5.5, fig.align = "center"}
# Compare minimal and optional reporting standards within databases
# MetaboLights
MLMin <- rbind(MLHumanCellP, MLHumanClinP, MLMouseP)
MLO <- rbind(MLHumanCellOP, MLHumanClinOP, MLMouseOP)
MLMin$Req <- rep("Minimal")
MLO$Req <- rep("Optional")
MLComp <- rbind(MLMin, MLO)
MLComp$Order <- as.factor(MLComp$Order)
# Plot box and Whisker diagram
ggplot(MLComp, aes(x=Order, y=Percentage, fill = Req))  + 
  geom_boxplot(outlier.shape = 3, outlier.color = "red")  +
  scale_x_discrete(labels = c(expression(italic("In vitro")),
                              "Clinical",
                              "Pre-clinical",
                              "Plant")) +
  theme_bw() +
  theme(
    axis.text = element_text(colour = "black"),
    axis.line.x = element_line(color="black", size = 0.5),
    axis.line.y = element_line(color="black", size = 0.5),
    # Remove gridlines and borders
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    legend.title = element_blank(),
    legend.position="top") +
  scale_y_continuous(expand = c(0, 0), limits = c(-1, 100)) +
  xlab("MSI Guideline") +
  ylab("Compliance (%)") +
  scale_fill_manual(values = c("#7fc97f", "#a6cee3"))
```

## Figure 2b.Comparison of Compliance to Minimal and Best Practice Reporting Standards in Metabolomics Workbench
Box-and-whisker plot showing the percentage compliance to the minimal and optional/best practice reporting standards in Metabolomics Workbench. Minimal reporting standards are indicated in purple, optional/best practice reporting standards are pink and red ‘+’ indicate outliers.

```{r MWCom, results = 'asis', fig.height = 4.5, fig.width = 5.5, fig.align = "center"}
# Compare minimal and optional reporting standards within databases
# MetaboLights
MWMin <- rbind(MWHumanCellP, MWHumanClinP, MWMouseP)
MWO <- rbind(MWHumanCellOP, MWHumanClinOP, MWMouseOP)
MWMin$Req <- rep("Minimal")
MWO$Req <- rep("Optional")
MWComp <- rbind(MWMin, MWO)
MWComp$Order <- as.factor(MWComp$Order)
# Plot box and Whisker diagram
ggplot(MWComp, aes(x=Order, y=Percentage, fill = Req))  + 
  geom_boxplot(outlier.shape = 3, outlier.color = "red")  +
  scale_x_discrete(labels = c(expression(italic("In vitro")),
                              "Clinical",
                              "Pre-clinical",
                              "Plant")) +
  theme_bw() +
  theme(
    axis.text = element_text(colour = "black"),
    axis.line.x = element_line(color="black", size = 0.5),
    axis.line.y = element_line(color="black", size = 0.5),
    # Remove gridlines and borders
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    legend.title = element_blank(),
    legend.position="top") +
  scale_y_continuous(expand = c(0, 0), limits = c(-1, 100)) +
  xlab("MSI Guideline") +
  ylab("Compliance (%)") +
  scale_fill_manual(values = c("#beaed4", "#fb9a99"))
```

## Figure 3. Compliance to Plant MSI Reporting Standards. 
Combined box-and-whisker and dot plots showing the percentage compliance of *A. thaliana* studies to the plant MSI minimum reporting standards within the MeRy-B, MetaboLights, Metabolomics Workbench and MetaPhen repositories. Red ‘+’ indicate outliers; each “dot” represents compliance to a single reporting standard. Letters denote significant differences in compliance (Kruskal Wallis, Dunn post-hoc test with Benjamini-Hochberg correction).
```{r plantfig, results = 'asis', fig.height = 4.5, fig.width = 4, fig.align = "center"}
MLPlantP$Repository <- rep("MetaboLights")
MWPlantP$Repository <- rep("Metabolomics Workbench")
MPPlantP$Repository <- rep("MetaPhen")
MeRPlantP$Repository <- rep("MeRy-B")
Plant <- rbind(MLPlantP, MWPlantP, MPPlantP, MeRPlantP)
Plant$Repository <- as.factor(Plant$Repository)
levels(Plant$Repository) <- gsub(" ", "\n", levels(Plant$Repository))
ggplot(Plant, aes(x=Repository, y=Percentage, fill=Repository))  + 
  geom_boxplot(outlier.shape = 3, outlier.color = "red")  +
  geom_jitter() +
  theme_bw() +
  theme(
    axis.text = element_text(colour = "black"),
    axis.line.x = element_line(color="black", size = 0.5),
    axis.line.y = element_line(color="black", size = 0.5),
    # Remove gridlines and borders
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    legend.position="none") +
  scale_y_continuous(expand = c(0, 0), limits = c(-1, 100)) +
  ylab("Compliance (%)") +
  scale_fill_manual(values = c("#fdc086","#7fc97f", "#beaed4", "#ffff99"))
```


## Supplementary Figure 1a. Compliance to Best Practice MSI Reporting Standards in MetaboLights
Combined box-and-whisker and dot plot showing the percentage compliance with the MSI best practice/optional reporting standards within the MetaboLights repository. Red `+' indicate outliers. 

```{r MLOp, results = 'asis', fig.height = 4.5, fig.width = 3.5, fig.align = "center"}
MLO <- rbind(MLHumanCellOP, MLHumanClinOP, MLMouseOP)
MLO$Order <- as.factor(MLO$Order)
ggplot(MLO, aes(x=Order, y=Percentage))  + 
  geom_boxplot(fill = "#7fc97f", outlier.shape = 3, outlier.color = "red")  +
  geom_jitter() +
  scale_x_discrete(labels = c(expression(italic("In vitro")),
                              "Clinical",
                              "Pre-clinical",
                              "Plant")) +
  scale_y_continuous(expand = c(0, 0), limits = c(-1, 101)) +
  theme_adjbw() +
  xlab("MSI Guideline") +
  ylab("Compliance (%)")
```

## Supplementary Figure 1b. Compliance to Best Practice MSI Reporting Standards in Metabolomics Workbench
Combined box-and-whisker and dot plot showing the percentage compliance with the MSI best practice/optional reporting standards within the Metabolomics Workbench Repository. Red `+' indicate outliers.

```{r MWOp, results = 'asis', fig.height = 4.5, fig.width = 3.5, fig.align = "center"}
MWO <- rbind(MWHumanCellOP, MWHumanClinOP, MWMouseOP)
MWO$Order <- as.factor(MWO$Order)
ggplot(MWO, aes(x=Order, y=Percentage))  + 
  geom_boxplot(fill = "#beaed4", outlier.shape = 3, outlier.color = "red")  +
  geom_jitter() +
  scale_x_discrete(labels = c(expression(italic("In vitro")),
                              "Clinical",
                              "Pre-clinical",
                              "Plant")) +
  theme_adjbw() +
  scale_y_continuous(expand = c(0, 0), limits = c(-1, 101)) +
  xlab("MSI Guideline") +
  ylab("Compliance (%)")
```

## Supplementary Figure 2. Species per Repository

The frequency of studies including different species in the metabolomics data repositories: MeRy-B, MetaboLights, Metabolomics Workbench and MetaPhen. For a species to be plotted as an individual band it must have been found in a minimum of nine studies across the repositories. Species found in less than nine studies across the four repositories are reported as Other Species.
```{r SpeciesperRep, results = 'asis', fig.height = 4.5, fig.width = 6, fig.align = "center"}
# Load csv
SpeciesbyRepository <- read.csv("../SpeciesperRepository.csv")
# Replace all spaces with line graps for a cleaner plot
levels(SpeciesbyRepository$Repository) <- gsub(" ", "\n", levels(SpeciesbyRepository$Repository))
ggplot(SpeciesbyRepository, aes(x=Repository, y=Frequency, fill=Species))  +
  geom_bar(colour="black", stat="identity", width = 0.7) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 400)) +
  scale_fill_manual(values = c('#80b1d3','#bebada','#b3de69',"#8dd3c7",
                                          '#fb8072',           
                                          '#ffffb3',
                                          '#fdb462',
                                          '#fccde5',
                                          '#d9d9d9'),
                               labels = c(expression(italic("Arabidopsis thaliana")), 
                                          expression(italic("Escherichia coli")), 
                                          expression(italic("Homo sapiens")),
                                          expression(italic("Lycopersicon esculentum")),
                                          expression(italic("Mus musculus")),
                                          expression(italic("Oryza sativas")),
                                          "Other Species",
                                          expression(italic("Rattus norvegicus")),
                                          expression(italic("Solanum lycopersicum"))))  +
  ylab("Number of Studies") +
  theme_bw() +
  theme(axis.text = element_text(colour = "black"),
        legend.text.align = 0,
        #legend.text = element_text(face = "italic"),
        axis.line.x = element_line(color="black", size = 0.5),
        axis.line.y = element_line(color="black", size = 0.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 
```

# Statistical Analysis
For analysing differences within repositories (ML and MW) Kruskal Wallis tests, followed by Dunn post-hoc tests with Benjamini-Hochberg correction were used. A Kruskal Wallis test was also used for comparing differences between repositories, both between ML and MW and between all four repositories for the plant guidelines. Mann-Whitney U tests were used to analyse difference between minimal and best practice reporting standards. 

```{r stats}
# Between MetaboLights minimal
KruskalML <- kruskal.test(Percentage ~ Order, data = ML)
DunnML <- dunnTest(Percentage ~ Order, data = ML, method = "bh") 

# Between Metabolomics Workbench minimal
KruskalMW <- kruskal.test(Percentage ~ Order, data = MW)
DunnMW <- dunnTest(Percentage ~ Order, data = MW, method = "bh")

# Between MetaboLights best practice/optional
KruskalMLO <- kruskal.test(Percentage ~ Order, data = MLO)
DunnMLO <- dunnTest(Percentage ~ Order, data = MLO, method = "bh") 

# Between Metabolomics Workbench best practice/optional
KruskalMWO <- kruskal.test(Percentage ~ Order, data = MWO)
DunnMWO <- dunnTest(Percentage ~ Order, data = MWO, method = "bh")                   

# Between repositories for plant guidelines only
# Add repostitory to plant data
KruskalPlant <- kruskal.test(Percentage ~ Repository, data = Plant)
DunnPlant <- dunnTest(Percentage ~ Repository, data = Plant, method = "bh")

# Compare Between repositories - MetaboLights and Metabolomics Workbench
# Kruskal Wallis - Minimal
ML$Repository <- rep("MetaboLights")
MW$Repository <- rep("Metablomics Workbench")
MLMW <- rbind(ML, MW)
MLMW$Repository <- as.factor(MLMW$Repository)
KruskalMLMW <- kruskal.test(Percentage ~ Repository, data = MLMW)

# Kruskal Wallis - Optional
MLO$Repository <- rep("MetaboLights")
MWO$Repository <- rep("Metablomics Workbench")
MLMWO <- rbind(MLO, MWO)
MLMWO$Repository <- as.factor(MLMWO$Repository)
KruskalMLMWO <- kruskal.test(Percentage ~ Repository, data = MLMWO)
                          
# Between Minimum and Optional
# MetaboLights
# Human - Cell Line
MLHumanCellP$MinBP <- rep("Min")
MLHumanCellOP$MinBP <- rep("BP")
MLHumanCellCom <- rbind(MLHumanCellP, MLHumanCellOP)
MLHumanCellCom$MinBP <- as.factor(MLHumanCellCom$MinBP)               
MannWhitMLHumanCell <- wilcox.test(Percentage ~ MinBP, data =  MLHumanCellCom)
# Human - Clinical 
MLHumanClinP$MinBP <- rep("Min")
MLHumanClinOP$MinBP <- rep("BP")
MLHumanClinCom <- rbind(MLHumanClinP, MLHumanClinOP)
MLHumanClinCom$MinBP <- as.factor(MLHumanClinCom$MinBP)               
MannWhitMLHumanClin <- wilcox.test(Percentage ~ MinBP, data =  MLHumanClinCom)
# Mouse - Pre-clinical
MLMouseP$MinBP <- rep("Min")
MLMouseOP$MinBP <- rep("BP")
MLMouseCom <- rbind(MLMouseP, MLMouseOP)
MLMouseCom$MinBP <- as.factor(MLMouseCom$MinBP)               
MannWhitMLMouse <- wilcox.test(Percentage ~ MinBP, data =  MLMouseCom)

# Metabolomics Workbench
# Human - Cell Line
MWHumanCellP$MinBP <- rep("Min")
MWHumanCellOP$MinBP <- rep("BP")
MWHumanCellCom <- rbind(MWHumanCellP, MWHumanCellOP)
MWHumanCellCom$MinBP <- as.factor(MWHumanCellCom$MinBP)               
MannWhitMWHumanCell <- wilcox.test(Percentage ~ MinBP, data =  MWHumanCellCom)
# Human - Clinical 
MWHumanClinP$MinBP <- rep("Min")
MWHumanClinOP$MinBP <- rep("BP")
MWHumanClinCom <- rbind(MWHumanClinP, MWHumanClinOP)
MWHumanClinCom$MinBP <- as.factor(MWHumanClinCom$MinBP)               
MannWhitMWHumanClin <- wilcox.test(Percentage ~ MinBP, data =  MWHumanClinCom)
# Mouse - Pre-clinical
MWMouseP$MinBP <- rep("Min")
MWMouseOP$MinBP <- rep("BP")
MWMouseCom <- rbind(MWMouseP, MWMouseOP)
MWMouseCom$MinBP <- as.factor(MWMouseCom$MinBP)               
MannWhitMWMouse <- wilcox.test(Percentage ~ MinBP, data =  MWMouseCom)
```

# Supplementary Tables

## Supplementary Table 1. Compliance of Human Clinical Studies to Minimal MSI Reporting Standards 

```{r humanclinmin, results = 'asis'}
HumanClinP <- cbind(MLHumanClinP[,1:2], MWHumanClinP[,2])
rownames(HumanClinP) <- HumanClinP$Metadata
colnames(HumanClinP) <- c("Metadata", "MetaboLights", "Metabolomics Workbench")
HumanClinP$total <- rowSums(HumanClinP[2:3])
HumanClinP <- HumanClinP[order(-HumanClinP$total),]
kable(HumanClinP[,2:3], caption = "The percentage of *Homo sapiens* studies in each repository that comply with each mammalian clinical trials and human studies minimal reporting standard. ")
```

## Supplementary Table 2. Compliance of Human Cell Line Studies to Minimal MSI Reporting Standards

```{r humancellmin, results = 'asis'}
HumanCellP <- cbind(MLHumanCellP[,1:2], MWHumanCellP[,2])
rownames(HumanCellP) <- HumanCellP$Metadata
colnames(HumanCellP) <- c("Metadata", "MetaboLights", "Metabolomics Workbench")
HumanCellP$total <- rowSums(HumanCellP[2:3])
HumanCellP <- HumanCellP[order(-HumanCellP$total),]
kable(HumanCellP[,2:3], caption = "The percentage of *Homo sapiens* studies in each repository that comply with each microbial and *in vitro* minimal reporting standard.")
```

## Supplementary Table 3. Compliance of Mouse Pre-Clinical Studies to Minimal MSI Reporting Standards

```{r mousemin, results = 'asis'}
MouseP <- cbind(MLMouseP[,1:2], MWMouseP[,2])
rownames(MouseP) <- MouseP$Metadata
colnames(MouseP) <- c("Metadata", "MetaboLights", "Metabolomics Workbench")
MouseP$total <- rowSums(MouseP[2:3])
MouseP <- MouseP[order(-MouseP$total),]
kable(MouseP[,2:3], caption = "The percentage of *Mus musculus* studies in each repository that comply with each pre-clinical minimal reporting standard.")
```

## Supplementary Table 4. Compliance of Arabidopis Studies to Minimal MSI Reporting Standards

```{r plantmin, results = 'asis'}
PlantP <- cbind(MeRPlantP[,1:2], MLPlantP[,2], MWPlantP[,2], MPPlantP[,2])
colnames(PlantP) <- c("Metadata", "MeRy-B", "MetaboLights", "Metabolomics Workbench", "MetaPhen")
PlantP$total <- rowSums(PlantP[2:5])
PlantP <- PlantP[order(-PlantP$total),]
kable(PlantP[,2:5], caption = "The percentage of *Arabidopsis thaliana* studies in each repository that comply with each plant minimal reporting standard.")
```

## Supplementary Table 5. Compliance of Human Clinical Studies to Best Practice MSI Reporting Standards   

```{r humanclinop, results = 'asis'}
HumanClinOP <- cbind(MLHumanClinOP[,1:2], MWHumanClinOP[,2])
rownames(HumanClinOP) <- HumanClinOP$Metadata
colnames(HumanClinOP) <- c("Metadata", "MetaboLights", "Metabolomics Workbench")
HumanClinOP$total <- rowSums(HumanClinOP[2:3])
HumanClinOP <- HumanClinOP[order(-HumanClinOP$total),]
kable(HumanClinOP[,2:3], caption = "The percentage of *Homo sapiens* studies in each repository that comply with each mammalian clinical trials and human studies optional reporting standard.")
```

## Supplementary Table 6. Compliance of Human Cell Line Studies to Best Practice MSI Reporting Standards  

```{r humancellop, results = 'asis'}
HumanCellOP <- cbind(MLHumanCellOP[,1:2], MWHumanCellOP[,2])
rownames(HumanCellOP) <- HumanCellOP$Metadata
colnames(HumanCellOP) <- c("Metadata", "MetaboLights", "Metabolomics Workbench")
HumanCellOP$total <- rowSums(HumanCellOP[2:3])
HumanCellOP <- HumanCellOP[order(-HumanCellOP$total),]
kable(HumanCellOP[,2:3], caption = "The percentage of *Homo sapiens* studies in each repository that comply with each microbial and *in vitro* best practice reporting standard")
```

## Supplementary Table 7. Compliance of Mouse Pre-Clinical Studies to Best Practice MSI Reporting Standards 

```{r mouseop, results = 'asis'}
MouseOP <- cbind(MLMouseOP[,1:2], MWMouseOP[,2])
rownames(MouseOP) <- MouseOP$Metadata
colnames(MouseOP) <- c("Metadata", "MetaboLights", "Metabolomics Workbench")
MouseOP$total <- rowSums(MouseOP[2:3])
MouseOP <- MouseOP[order(-MouseOP$total),]
kable(MouseOP[,2:3], caption = "The percentage of *Mus musculus* studies in each repository that comply with each pre-clinical optional reporting standard.")
```

